The demographic costs of sexually antagonistic selection in partially selfing populations

### 2021-12-31 doc/SA-Hermaphrodites-wDemography-R1.tex
Minor tweaks to intro
Realized I didn't like that last sentence about eco-evo, so edited.

%To persist in the long term, a population must be able to adapt to its environment. Yet, the process of adaptation can be impeded by a variety of genetic and environmental factors, such as deleterious mutations \citep{Haldane1937}, \textcolor{blue}{rapid, unpredictable, or extreme changes} in environmental conditions \citep{Maynard-Smith1976, LandeShannon1996, OrrUnckless2008,Lyberger-etal-2021}, maladaptive gene-flow \citep{KirkpatrickBarton1997, BolnickNosil2007}, and genetic constraints \citep{ConnallonHall2018, MatthewsConnallon2019}. Genetic constraints on adaptation can arise as a consequence of conflicting selection and gene-flow between different classes of individuals, or more generally, when \textcolor{blue}{different fitness components are negatively correlated} \citep{CharlesworthHughes2000, ConnallonHall2018}. 
\textcolor{blue}{A population's ability to adapt to its environment and persist in the long term depends upon the nature of the genetic variation that it harbors \citep{Fisher1930, Lewontin1975}. Yet, species exist in a complex world where the process of adaptation can be impeded by conflicting selection and gene-flow between different classes of individuals, or more generally, when different fitness components are negatively correlated \citep{CharlesworthHughes2000, ConnallonHall2018}.}

@@

Sexually antagonistic selection (abbreviated SA hereafter) \textcolor{blue}{is an important form of conflicting selection that occurs when beneficial alleles for one sex are deleterious when expressed in the other. Likewise, alleles with opposing fitness effects through male and female sex functions can cause SA selection in hermaphrodite populations, where individuals' fitness is determined jointly by their maternal and paternal reproductive success \citep{LloydWebb1986, WebbLloyd1986, Abbott2011, JordanConnallon2014}. SA selection appears to be} a common feature of sexually reproducing populations, and is thought to contribute to the maintenance of genetic variation and the evolution of sexual dimorphism and sex-related traits \citep{Kidwell1977, Lande1980, Rice1992, Charlesworth1999, RiceChippindale2001, ConnallonClark2012, BondurianskyChenoweth2009, Olito2019}. \textcolor{blue}{For instance, phenotypic selection studies commonly find sex-differences in directional selection \citep[e.g.,][]{DeLisle-etal-2018,Lewis-etal-2011,CoxCalsbeek2009,SinghPunzalan2018}, implying that many genetic variants underlying quantitative traits have SA fitness effects. Likewise, quantitative genetic studies frequently find negative cross-sex genetic correlations for fitness (\citealt{Chippendale-etal-2001,Delph-etal-2011}; reviewed in \citealt{ConnallonMatthews2019}), and polygenic signals of SA selection are evident in recent population genetic analyses of large genomic datasets \citep{RuzickaESEB2020, RuzickaConnallon2022}. Overall, both theory and current empirical data suggest that there is ample scope for the maintenance of SA \textcolor{blue}{genetic variation} in both dioecious \textcolor{blue}{(separate-sexed)} and hermaphrodite populations \citep{BondurianskyChenoweth2009, Abbott2011, RuzickaESEB2020}.}

@@

\textcolor{blue}{While significant efforts have been made to predict when SA polymorphisms should be maintained and to detect SA genetic variation from empirical data, less attention has been paid to the demographic consequences of SA selection \citep{MatthewsConnallon2019}}. Most population genetic theory, including SA selection theory, models the relative fitness of genotypes in populations of constant (and often infinite) size. In reality, these genotypic fitnesses emerge from myriad processes acting throughout the life-history of individuals \textcolor{blue}{which link genotype distributions to population dynamics (e.g., \citealt {johnston2013life, merot2020balancing}). For over 50 years, the field of eco-evolutionary dynamics has sought to jointly model the genetic and demographic processes involved in evolutionary change. Theoretical models combining population genetics with density-dependent population growth \citep{roughgarden1971density} and stage-structured populations (consolidated in \citealt{Charlesworth_1994}) paved the way toward understanding the demographic consequences of allele frequency changes due to selection. Various extensions of these theories have been made to study life-history evolution \citep{orive1995senescence, orive2001somatic}, spatial structure \citep{ronce2000kin}, kin selection \citep{rousset2004genetic}, and the evolution of quantitative traits \citep[e.g.,][]{coulson2006putting, barfield2011evolution, childs2016evolution, orive2017effects}. Together, these eco-evolutionary models highlight that a fuller understanding of the genetic basis of fitness variation, and the processes maintaining it, requries an integration of genetic and demographic evolutionary processes.}

>>>>

%To persist in the long term, a population must be able to adapt to its environment. Yet, the process of adaptation can be impeded by a variety of genetic and environmental factors, such as deleterious mutations \citep{Haldane1937}, \textcolor{blue}{rapid, unpredictable, or extreme changes} in environmental conditions \citep{Maynard-Smith1976, LandeShannon1996, OrrUnckless2008,Lyberger-etal-2021}, maladaptive gene-flow \citep{KirkpatrickBarton1997, BolnickNosil2007}, and genetic constraints \citep{ConnallonHall2018, MatthewsConnallon2019}. Genetic constraints on adaptation can arise as a consequence of conflicting selection and gene-flow between different classes of individuals, or more generally, when \textcolor{blue}{different fitness components are negatively correlated} \citep{CharlesworthHughes2000, ConnallonHall2018}. 
\textcolor{blue}{A population's ability to adapt to its environment and persist in the long term depends upon the nature of the genetic variation that it harbors \citep{Fisher1930, Lewontin1975}. Yet, species exist in a complex world where the process of adaptation can be impeded by conflicting selection and gene-flow between different classes of individuals, or more generally, when different fitness components are negatively correlated \citep{Prout2000, CharlesworthHughes2000, ConnallonHall2018}.}

@@

Sexually antagonistic selection (abbreviated SA hereafter) \textcolor{blue}{is an important form of conflicting selection that occurs when beneficial alleles for one sex are deleterious when expressed in the other. Likewise, alleles with opposing fitness effects through male and female sex functions can cause SA selection in hermaphrodite populations, where individuals' fitness is determined jointly by their maternal and paternal reproductive success \citep{LloydWebb1986, WebbLloyd1986, Abbott2011, JordanConnallon2014}. SA selection appears to be} a common feature of sexually reproducing populations, and is thought to contribute to the maintenance of genetic variation and the evolution of sexual dimorphism and sex-related traits \citep{Kidwell1977, Lande1980, Rice1992, Charlesworth1999, RiceChippindale2001, ConnallonClark2012, BondurianskyChenoweth2009, Olito2019}. \textcolor{blue}{For instance, phenotypic selection studies commonly find sex-differences in directional selection \citep[e.g.,][]{DeLisle-etal-2018,Lewis-etal-2011,CoxCalsbeek2009,SinghPunzalan2018}, implying that many genetic variants underlying quantitative traits have SA fitness effects. Likewise, quantitative genetic studies frequently find negative cross-sex genetic correlations for fitness (\citealt{Chippendale-etal-2001,Delph-etal-2011}; reviewed in \citealt{ConnallonMatthews2019}), and polygenic signals of SA selection are evident in recent population genetic analyses of large genomic datasets \citep{RuzickaESEB2020, RuzickaConnallon2022}. Overall, both theory and current empirical data suggest that there is ample scope for the maintenance of SA \textcolor{blue}{genetic variation} in both dioecious \textcolor{blue}{(separate-sexed)} and hermaphrodite populations \citep{BondurianskyChenoweth2009, Abbott2011, RuzickaESEB2020, WangBarrett2020}.}

@@

\textcolor{blue}{While significant efforts have been made to predict when SA polymorphisms should be maintained and to detect SA genetic variation from empirical data, less attention has been paid to the demographic consequences of SA selection \citep{MatthewsConnallon2019}}. Most population genetic theory, including SA selection theory, models the relative fitness of genotypes in populations of constant (and often infinite) size. In reality, these genotypic fitnesses emerge from myriad processes acting throughout the life-history of individuals \textcolor{blue}{which link genotype distributions to population dynamics (e.g., \citealt {johnston2013life, merot2020balancing}). For over 50 years, the field of eco-evolutionary dynamics has sought to jointly model the genetic and demographic processes involved in evolutionary change. Theoretical models combining population genetics with density-dependent population growth \citep{roughgarden1971density} and stage-structured populations (consolidated in \citealt{Charlesworth_1994}) paved the way toward understanding the demographic consequences of allele frequency changes due to selection. Various extensions of these theories have been made to study life-history evolution \citep{orive1995senescence, orive2001somatic}, spatial structure \citep{ronce2000kin}, kin selection \citep{rousset2004genetic}, and the evolution of quantitative traits \citep[e.g.,][]{coulson2006putting, barfield2011evolution, childs2016evolution, orive2017effects}. This growing body of eco-evolutionary theory highlights that demographic processes must be taken into account explicitly in order to fully understand how genetic variation for fitness is shaped and maintained in natural populations.}
